Rank Verification for Exponential Families

### 2016-06-30 verifying-winner.tex
more comments / small edits

\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage[totalwidth=480pt, totalheight=680pt]{geometry}
\usepackage{mathrsfs}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{xcolor}
\usepackage[round]{natbib}
\definecolor{mid_green}{HTML}{99d8c9}

@@

Following the framework in \citet{Fithian:2014ws}, we consider the selection event where the winner is $1$: 
$$A_1 = \left\{X_1 \ge \max_{j > 1} X_j\right\}.$$
For convenience, we let
$$D_{jk} = \frac{X_j - X_k}{2} ~~~~ \text{ and } ~~~~ M_{jk} = \frac{X_j + X_k}{2}.$$

@@

Without loss of generality, it is sufficient to show that $p_{12} \ge p_{13}$. In the computation of either $p$-value, $X_4, \ldots, X_n$ are all conditioned on. Fixing these variables that are conditioned on, $g$ has only thee variables: $X_1$, $X_2$ and $X_3$. We can write this as $g_{X_4, \ldots, X_n}\left(X_1, X_2, X_3\right)$. The first point in \Cref{lma:twoprop} says that the function $g_{X_4, \ldots, X_n}$ is Schur-concave as well. This allows us to clean up our notation and just rename $g_{X_4, \ldots, X_n}$ as $g$. The p-values thus are
\begin{align*}
p_{12} & = \frac{\int_{D_{12}}^\infty g\left(M_{12} + z, M_{12} - z, X_3\right) \,dz}{\int_0^\infty g\left(M_{12} + z, M_{12} - z, X_3\right) \,dz} \\
p_{13} & = \frac{\int_{D_{13}}^\infty g\left(M_{13} + z, X_2, M_{13} - z\right) \,dz}{\int_{\max\left\{X_2 - M_{13}, 0\right\}}^\infty g\left(M_{13} + z, X_2, M_{13} - z\right) \,dz}
\end{align*}

@@

@@

A fundamental difference between the two methods is the definition of the threshold. The threshold from \citet{Gupta:1967wg} depends solely on the setup of the model: $m$ and $n$. Our threshold, however, depends on the data itself. Such difference is amplified when $n$ is large. The threshold from \citeauthor{Gupta:1967wg} increases with $n$ even if the added observations are all $0$. In such case, our threshold does not change at all, suggesting more potential test power.

@@

We can take the idea above further and construct confidence bound for $\theta_{[1]} - \max_{j \ne [1]} \theta_{j}$. This can be achieved by considering a statistical test, to see if $\theta_{[1]} - \max_{j \ne [1]} \theta_{j} \le \delta$, followed by the duality of confidence interval and statistical test.

Here we provide a more powerful Procedure 2' first. The confidence lower bound can be taken as the smallest $\delta$ such that we can reject $H_{0[1]}^\delta: \theta_{[1]} - \max_{j \ne [1]} \theta_{[j]} > \delta$. Once again this can be written as a union of null hypothesis
$$H_{0[1]}^\delta = \bigcup_{j \ne [1]} H_{0[1]j}: \theta_1 - \theta_j > \delta.$$
By \Cref{lma:union}, we can construct selective exact one-tailed $p$-values $p_{[1]j}$ for each of these by conditioning on $A_{[1]}$, $M_{[1]j}$ and $X_{\setminus\left\{[1], j\right\}}$, giving us an exact test for $H_{0[1]}$ by rejecting whenever $\max_{j \ne [1]} p_{[1]j} < \alpha$.

\begin{theorem}
The $p$-values constructed above satisfies $p_{[1][2]} \ge p_{[1]j}$ for any $j \ne [1]$.
\end{theorem}

\begin{proof}
Again we start with assuming $X_1 \ge \cdots \ge X_n$. The $p$-values in question are derived from the conditional law
$$\mathcal{L}_{\theta_1 - \theta_j = \delta} \left(D_{1j} \middle| D_{1j}, X_2, \ldots, X_n, A\right),$$
which is the truncated distribution
\begin{align*}
p\left(D_{1j}\right) & \propto \exp\left(\left(\theta_1 - \theta_j\right) D_{1j} + \theta_2 X_2 + \cdots + \left(\theta_1 + \theta_j\right) M_{1j} + \cdots + \theta_n X_n \right) \\
& ~~~~~~~~ g\left(M_{1j} + D_{1j}, X_2, \ldots, M_{1j} - D_{1j}, \ldots X_n\right) 1_{A_1} \\
& \propto \exp\left(\delta D_{1j}\right) g\left(M_{1j} + D_{1j}, X_2, \ldots, M_{1j} - D_{1j}, \ldots X_n\right) 1_{A_1}.
\end{align*}

@@

In other words, Procedure 2' can be summarized as: Find the minimum $\delta$ such that $p_{12} \le \alpha$. And by construction, Procedure 2' gives exact $1-\alpha$ confidence bound for $\theta_{[1]} - \max_{j \ne [1]} \theta_j$.

@@

\begin{proof}
If the confidence bound $\delta^*$ is negative, we report $-\infty$ so it definitely is valid and conservative. On the other hand, if $\delta^* \ne 0$, then at $\delta^*$ the $p$-value $p_{12} = \alpha$. The denominator is the positive part of a positively skewed distribution, and thus bigger than half of the distribution. So the unadjusted pairwise test $p$-value from Procedure 2 comparing $X_1$ and $X_2$ is larger than $p_{12}$ from Procedure 2', rendering Procedure 2 conservative.
\end{proof}

Note that Procedure 2 reporting $-\infty$ in case of $\delta^*$ is rather extreme. In reality, we can always just adopt Procedure 2' in the case when Procedure 1 rejects. In fact, by Procedure 2', the multinomial example for polling in \Cref{sec:iowa} can give a stronger lower confidence bound, that $\log \pi_{\text{Trump}} - \max_{j \ne \text{Trump}} \log \pi_{j} \ge 0.103142$.

@@

\begin{theorem}
There is a step-down procedure, which we will denote as Procedure 3', that infers $j_0$ above with the FWER controlled at $\alpha$.
\end{theorem}

@@

This can be fitted into the selective sequential model selection framework proposed by \citet{Fithian:2015uj}. Given the observations $X_1 \ge \cdots \ge X_n$, drawn from an unknown distribution $F$, we consider the sequence of null hypothesis of
$$H_{0j}: \theta_j \le \max_{k > j} \theta_k.$$
Alternatively, we can think of them as a sequence of nested model
$$M_0\left(X\right) \subseteq M_1\left(X\right) \subseteq \cdots \subseteq M_n\left(X\right), ~~~~~~~~ M_j\left(X\right)^c = \left\{\theta_1 > \cdots > \theta_j > \max_{k > j} \theta_k\right\},$$
and the each $M_j$ depends on $X$ through the order of $X$. The step-down procedure for rejecting $H_{0j}$ is equivalent to selecting the smallest $j$ for which $F \in M_j\left(X\right)$.

@@

Here we are only comparing $X_j$ to $X_{j+1}$ by \Cref{sec:winnerbest}. The upper truncation for $X_j$ can be represented by cropping \Cref{fig:compare_rays} along a vertical line, shown in \Cref{fig:crop_rays}, thus the proof of Part 1 of \Cref{thm:main} on sufficiency to compare $X_1$ and $X_2$ remains valid with appropriate conditioning. Again we can construct the $p$-values as in \Cref{eqn:p1j} for $k>j$.

@@

Hence similar to the proof of Part 1 of \Cref{thm:main}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in \citet{Fithian:2015uj}, as
$$\PP_F \left(p_j \le \alpha \middle| M_{j-1}\left(X\right)\right) \le \alpha \text{ a.s.}, ~~~~~~~~ F \in M_{j-1}\left(X\right)$$
by tower property.

@@

Our selective test procedure allows us to untangle multiple comparisons based on correlated observations easily, while avoiding the pitfalls of selection. To illustrate, we look at a past psychology study done by \citet{Uhls:2012gf}.

\begin{table}[htbp]
\begin{center}
\begin{tabular}{c c}
	\hline
	Value & Count \\
	\hline
	Achievement & $3$ \\
	Benevolence & $5$ \\
	Fame & $8$ \\
	Community feeling & $2$ \\
	Financial success & $1$ \\
	Self-acceptance & $0$ \\
	Image & $0$ \\
	Other & $1$ \\
	\hline
\end{tabular}
\end{center}
\caption{The number of study participants who chose each value as their top desired one. Since only one option can be chosen, these observations are inherently multinomial.}
\label{tbl:fame}
\end{table}

A sample of $20$ students from elementary and middle schools were asked to choose a top ``value'' from 7 different options: achievement, benevolence, fame, community feeling, financial success, self-acceptance, image and other. The results are tabulated in \Cref{tbl:fame}. The study postulates that fame is significantly the top value of choice. It proceeds to perform a binomial test for the null hypothesis that the participants chose among the $7$ options uniformly and concluded with a $p$-value of $0.006$ that fame is indeed the top desired value. The analysis failed to address the ``winner's curse'', as well as the dependence among the observations. Our paper addressed these two issues and provided a simple test --- perform a unadjusted (two-tailed) binomial test of the winner and the runner-up. In this case, we test if $8$ is a significant observation in $\text{Binomial}\left(8 + 5, 0.5\right)$. The two-tailed $p$-value is $0.580811$, suggesting that the study was inconclusive.

This paper has provided a straightforward yet powerful method for verifying that the winner is the best, for joint distributions in the exponential family with mild condition. Such a test allows us to ``ignore'' the issues of selection and multiple-testing, which are often overlooked or mishandled.

>>>>

\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage[totalwidth=480pt, totalheight=680pt]{geometry}
\usepackage{mathrsfs}
\usepackage{verbatim}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{xcolor}
\usepackage[round]{natbib}
\definecolor{mid_green}{HTML}{99d8c9}

@@

Our proof has three main parts: (1) deriving $p_{1j}$ for each $j\geq 2$, (2) showing that $p_{12} \geq p_{1j}$ for each $j\geq 2$, and 
(3) showing that $p_{12}$ is an unadjusted pairwise $p$-value.

\paragraph{Derivation of $p_{1j}$}
Following the framework in \citet{Fithian:2014ws}, we consider the selection event where the winner is $1$: 
$$A_1 = \left\{X_1 \ge \max_{j > 1} X_j\right\}.$$
For convenience, we let
$$D_{jk} = \frac{X_j - X_k}{2} ~~~~ \text{ and } ~~~~ M_{jk} = \frac{X_j + X_k}{2}.$$

@@

\paragraph{Demonstration that $p_{1*}=p_{12}$}

We now proceed to show that $p_{12}$, the $p$-value comparing the winner to the runner-up, is the largest of all $p_{1j}$. Without loss of generality, it is sufficient to show that $p_{12} \ge p_{13}$. We can write this as $g_{X_4, \ldots, X_n}\left(X_1, X_2, X_3\right)$. The first point in \Cref{lma:twoprop} says that the function $g_{X_4, \ldots, X_n}$ is Schur-concave as well. This allows us to clean up our notation and just rename $g_{X_4, \ldots, X_n}$ as $g$. The p-values thus are
\begin{align*}
p_{12} & = \frac{\int_{D_{12}}^\infty g\left(M_{12} + z, M_{12} - z, X_3\right) \,dz}{\int_0^\infty g\left(M_{12} + z, M_{12} - z, X_3\right) \,dz} \\
p_{13} & = \frac{\int_{D_{13}}^\infty g\left(M_{13} + z, X_2, M_{13} - z\right) \,dz}{\int_{\max\left\{X_2 - M_{13}, 0\right\}}^\infty g\left(M_{13} + z, X_2, M_{13} - z\right) \,dz}
\end{align*}

@@

\paragraph{3. $p_{12}$ is an unadjusted pairwise $p$-value}

@@

A fundamental difference between the two methods is the definition of the threshold. The threshold from \citet{Gupta:1967wg} depends solely on the setup of the model: $m$ and $n$. Our threshold, however, depends on the data itself. Such difference is amplified when $n$ is large. The threshold from \citeauthor{Gupta:1967wg} increases with $n$ even if the added observations are all $0$. In such case, our threshold does not change at all, suggesting more potential test power.\WFcomment{Actually this isn't really why ours is more powerful.}

@@

By generalizing the above, we can construct a lower confidence bound for $\theta_{[1]} - \max_{j \ne [1]} \theta_{j}$. Here we provide a more powerful Procedure 2' first. We will proceed by inverting a statistical test of the hypothesis $H_{0[1]}^\delta:\; \theta_{[1]} - \max_{j \ne [1]} \theta_{j} \le \delta$, which can be written as a union of null hypotheses:
$$H_{0[1]}^\delta = \bigcup_{j \ne [1]} H_{0[1]j}: \theta_{[1]} - \theta_j \le \delta.$$
By \Cref{lma:union}, we can construct selective exact one-tailed $p$-values $p_{[1]j}^{(\delta)}$ \WFcomment{What do you think of attaching something like a superscript here to make it clear there are different $p$-values for every $\delta$?} for each of these by conditioning on $A_{[1]}$, $M_{[1]j}$ and $X_{\setminus\left\{[1], j\right\}}$, giving us an exact test for $H_{0[1]}$ by rejecting whenever $\max_{j \ne [1]} p_{[1]j}^{(\delta)} < \alpha$.

\begin{theorem}
The $p$-values constructed above satisfy $p_{[1][2]}^{(\delta)} \ge p_{[1]j}^{(\delta)}$ for any $j \ne [1]$.
\end{theorem}

\begin{proof}
Again we start with assuming $X_1 \ge \cdots \ge X_n$. The $p$-values in question are derived from the conditional law
$$\mathcal{L}_{\theta_1 - \theta_j = \delta} \left(D_{1j} \middle| M_{1j}, X_2, \ldots, X_n, A\right),$$
which is the truncated distribution
\begin{align*}
p\left(D_{1j}\right) & \propto \exp\left(\left(\theta_1 - \theta_j\right) D_{1j} + \theta_2 X_2 + \cdots + \left(\theta_1 + \theta_j\right) M_{1j} + \cdots + \theta_n X_n \right) \\
& ~~~~~~~~ g\left(M_{1j} + D_{1j}, X_2, \ldots, M_{1j} - D_{1j}, \ldots X_n\right) 1_{A_1} \\
& \propto \exp\left(\delta D_{1j}\right) g\left(M_{1j} + D_{1j}, X_2, \ldots, M_{1j} - D_{1j}, \ldots X_n\right) 1_{A_1}.
\end{align*}

@@

In other words, Procedure 2' can be summarized as: Find the minimum $\delta$ such that $p_{[1][2]}^{(\delta)} \le \alpha$. And by construction, Procedure 2' gives exact $1-\alpha$ confidence bound for $\theta_{[1]} - \max_{j \ne [1]} \theta_j$.

@@

\begin{proof}
If the confidence bound $\delta^*$ is negative, we report $-\infty$ so it definitely is valid and conservative. On the other hand, if $\delta^* \ne 0$, then at $\delta^*$ the $p$-value $p_{12} = \alpha$. The denominator is the positive part of a positively skewed distribution, and thus bigger than half of the distribution. So the unadjusted pairwise test $p$-value from Procedure 2 comparing $X_1$ and $X_2$ is larger than $p_{12}$ from Procedure 2', rendering Procedure 2 conservative.

\WFcomment{I found this proof confusing.}
\end{proof}

Note that Procedure 2 reporting $-\infty$ in case of $\delta^*$ is rather extreme. In reality, we can always just adopt Procedure 2' in the case when Procedure 1 rejects. In fact, by Procedure 2', the multinomial example for polling in \Cref{sec:iowa} can give a stronger lower confidence bound, that $\log \pi_{\text{Trump}} - \max_{j \ne \text{Trump}} \log \pi_{j} \ge 0.103$.\WFcomment{0.103142 is a lot of sig. digits.}%142$.

@@

\begin{theorem}
There is a step-down procedure, which we will denote as Procedure 3', that infers $j_0$ above with the FWER controlled at $\alpha$.
\end{theorem}
\WFcomment{Probably better to define Procedure 3' outside the theorem?}

@@

This can be fitted into the selective sequential model selection framework proposed by \citet{Fithian:2015uj}. Given the observations $X_1 \ge \cdots \ge X_n$, drawn from an unknown distribution $F$, we consider the following sequence of null hypotheses:
$$H_{0j}: \theta_j \le \max_{k > j} \theta_k.$$
Alternatively, we can think of them as a sequence of nested model
$$M_0\left(X\right) \subseteq M_1\left(X\right) \subseteq \cdots \subseteq M_n\left(X\right), ~~~~~~~~ M_j\left(X\right)^c = \left\{\theta_1 > \cdots > \theta_j > \max_{k > j} \theta_k\right\},$$
\WFcomment{You're overloading the variable name $M$ here}

and the each $M_j$ depends on $X$ through the order of $X$. The step-down procedure for rejecting $H_{0j}$ is equivalent to selecting the smallest $j$ for which $F \in M_j\left(X\right)$. \WFcomment{More like the smallest $j$ for which we cannot reject...}

@@

Here we are only comparing $X_j$ to $X_{j+1}$ by \Cref{sec:winnerbest}. The upper truncation for $X_j$ can be represented by cropping \Cref{fig:compare_rays} along a vertical line, shown in \Cref{fig:crop_rays}, thus the proof of Part 1 of \Cref{thm:main} on sufficiency\WFcomment{Careful about the word ``sufficiency'', it's a bit of a loaded term in stats} to compare $X_1$ and $X_2$ remains valid with appropriate conditioning. Again we can construct the $p$-values as in \Cref{eqn:p1j} for $k>j$.

@@

Hence similar to the proof of Part 1 of \Cref{thm:main}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in \citet{Fithian:2015uj}, as
$$\PP_\theta \left(p_j \le \alpha \mid \theta_{[i]} \text{ is not the } i\text{th largest, some } i\leq j\right) \le \alpha \text{ a.s.}, \quad \forall \; \theta$$
$$\PP_F \left(p_j \le \alpha \middle| M_{j-1}\left(X\right)\right) \le \alpha \text{ a.s.}, ~~~~~~~~ F \in M_{j-1}\left(X\right)$$
by tower property. \WFcomment{Careful on notation again for ``$F\in M_{j-1}$,'' the right hand side is random.}

@@

Our selective test procedure allows us to untangle multiple comparisons based on correlated observations easily, while avoiding the pitfalls of selection. Claims reporting the ``winner'' are commonly made in the scientific literature, usually with no significance level reported or an incorrect method applied. For example, \citet{Uhls:2012gf} asked $n=20$ elementary and middle school students which of seven personal values they most hoped to embody as adults: ``Fame'' (8 selected), ``Benevolence'' (5), ``Achievement'' (3), and several others (4 total including one response of ``Other''). The authors' main finding --- which appeared in the abstract, the first paragraph of the article, and later a CNN.com headline --- was that ``Fame'' was the most common response. This claim was accompanied by a signficance level of $0.006$, which the authors computed by testing whether the probability of selecting ``Fame'' was larger than 1/7.

\begin{comment}
\begin{table}[htbp]
\begin{center}
\begin{tabular}{c c}
	\hline
	Value & Count \\
	\hline
	Achievement & $3$ \\
	Benevolence & $5$ \\
	Fame & $8$ \\
	Community feeling & $2$ \\
	Financial success & $1$ \\
	Self-acceptance & $0$ \\
	Image & $0$ \\
	Other & $1$ \\
	\hline
\end{tabular}
\end{center}
\caption{The number of study participants who chose each value as their top desired one. Since only one option can be chosen, these observations are inherently multinomial.}
\label{tbl:fame}
\end{table}

A sample of $20$ students from elementary and middle schools were asked to choose a top ``value'' from 7 different options: achievement, benevolence, fame, community feeling, financial success, self-acceptance, image and other. The results are tabulated in \Cref{tbl:fame}. The study postulates that fame is significantly the top value of choice. It proceeds to perform a binomial test for the null hypothesis that the participants chose among the $7$ options uniformly and concluded with a $p$-value of $0.006$ that fame is indeed the top desired value. The analysis failed to address the ``winner's curse'', as well as the dependence among the observations.

We have provided a simple, powerful and straightforward test based on comparing the winner to the runner-up.

comparing the number of ``Fame'' responses to the distribution

 --- perform a unadjusted (two-tailed) binomial test of the winner and the runner-up. In this case, we test if $8$ is a significant observation in $\text{Binomial}\left(8 + 5, 0.5\right)$. The two-tailed $p$-value is $0.580811$, suggesting that the study was inconclusive.
\end{comment}

The obvious error in the authors' reasoning is perhaps understandable since there is, presently, no agreed-upon statistical solution for this commonplace problem. Using our approach, they could have performed a two-tailed binomial test of the winner (``Fame'') vs. the runner-up (``Benevolence''). This test returns a $p$-value of
\[
p = 2\PP\left[\text{Binomial}\left(8 + 5, 0.5\right) \geq 8\right] = 0.58,
\]
indicating that their data did not support their major finding.

This paper has provided a straightforward yet powerful method for verifying that the winner is the best, for joint distributions in the exponential family with a mild technical condition. Such a test allows us to ``ignore'' the issues of selection and multiple-testing, which are often overlooked or mishandled.\WFcomment{This last sentence might give the wrong impression.}
